Classical Planning Algorithms ¡ on the Atari Video Games

### 2014-12-01 icaps15-paper/paper.tex
Final AAAI Submission

We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
in comparison  with breadth-first search (\BRFS) and \UCT.
Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
The discount factor used by  all the algorithms is $\gamma=0.995$.
The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
 $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
The bound on the number of simulated frames is like a bound on
lookahead time, as most of the time in the lookahead  is spent
in calls to the emulator for computing the next RAM state. 
This is why the average time per action is similar to all the algorithms
except $\IW(1)$, that due to its aggressive pruning does not always use
the full budget and takes less time per action on average.

@@

Last, \IW(1) and \BFS\  in the lookahead select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. We'll look at this further in the future.

@@

Likewise, in \Freeway\ and \Berzerk\ both $\IW(1)$ and \BFS\ attain a better
score than the baseline semi-random algorithm \emph{Perturb} in
\cite{bellemare:jair2013}, that beats \UCT\ on those games.
%% to say the 2 lines below must add context: .. eg
%%% ``which in *these games* outperforms both \UCT\ and the reinforcement learning methods.''
%% add it if that's the case ..
\emph{Perturb} is a simple algorithm
that selects  a fixed  action with probability $0.95$, and a random
action with probability $0.05$.  For \emph{Perturb}, BNVB do not report 
the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
are deep but can be reached by  repeating the same action.
This is the case of \Freeway, where a chicken has to run to the top of the screen across a ten lane
highway filled with traffic. Every time the chicken gets across (starting at the bottom), there is one
unit of reward. If the chicken is hit by a car, it goes back some lanes. 
In \Freeway, only $12$ out of the $18$ possible actions have an effect:
$6$ actions move the chicken up (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
$6$ actions move the chicken down (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
when the selected fixed action moves the chicken up.  As noted 
in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
across the highway at all. 
%
The reason that \UCT\ does not collect any reward is that it needs to move the chicken up
at least 240 times\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
highway.} something that is very unlikely in a random exploration. $\IW(1)$ does not have this limitation and is best in \Freeway.
%% The reason for \UCT\ to perform erratically 
%% is that the roll-outs do not manage to collect any reward. A sequence of actions 
%% with at least $48$ actions embedded that move the chicken up are required in order 
%% to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
%% is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
%% that \emph{also} avoids incoming traffic in a timely fashion is rather low.
%% %% the previous number was not correct; no-op actions do not count, and thus needs
%% %%% 200 more move up actions than move down actions, although this is not accurate either
%% %%%% as move down actions that take chicken below the screen count as no-ops too ..
%% $\IW(1)$ does not have this limitation and is best in \Freeway.

@@

\noindent \textbf{Acknowledgments.} We thank the people involved in
the creation and maintenance of the Arcade Learning Environment that
have made this work possible. This research was co-funded by the
Australian Research Council linkage grant LP11010015.

>>>>

We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
in comparison  with breadth-first search (\BRFS) and \UCT.
Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
The discount factor used by  all the algorithms is $\gamma=0.995$.
The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
 $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
The bound on the number of simulated frames is like a bound on
lookahead time, as most of the time in the lookahead  is spent
in calls to the emulator for computing the next RAM state. 
This is why the average time per action is similar to all the algorithms
except $\IW(1)$, that due to its pruning does not always use
the full budget and takes less time per action on average.

@@

Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. We'll look at this further in the future.

@@

Likewise, in \Freeway\ and \Berzerk\ both $\IW(1)$ and \BFS\ attain a better
score than the baseline semi-random algorithm \emph{Perturb} in
\cite{bellemare:jair2013}, that beats \UCT\ on those games.
%% to say the 2 lines below must add context: .. eg
%%% ``which in *these games* outperforms both \UCT\ and the reinforcement learning methods.''
%% add it if that's the case ..
\emph{Perturb} is a simple algorithm
that selects  a fixed  action with probability $0.95$, and a random
action with probability $0.05$.  For \emph{Perturb}, BNVB do not report 
the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
are deep but can be reached by  repeating the same action.
This is the case of \Freeway, where a chicken has to run to the top of the screen across a ten lane
highway filled with traffic. Every time the chicken gets across (starting at the bottom), there is one
unit of reward. If the chicken is hit by a car, it goes back some lanes. 
In \Freeway, only $12$ out of the $18$ possible actions have an effect:
$6$ actions move the chicken up (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
$6$ actions move the chicken down (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
when the selected fixed action moves the chicken up.  As noted 
in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
across the highway at all. 
%
The reason that \UCT\ does not collect any reward is that it needs to move the chicken up
at least 240 times\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get it across the 
highway.} something that is very unlikely in a random exploration. $\IW(1)$ does not have this limitation and is best in \Freeway.
%% The reason for \UCT\ to perform erratically 
%% is that the roll-outs do not manage to collect any reward. A sequence of actions 
%% with at least $48$ actions embedded that move the chicken up are required in order 
%% to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
%% is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
%% that \emph{also} avoids incoming traffic in a timely fashion is rather low.
%% %% the previous number was not correct; no-op actions do not count, and thus needs
%% %%% 200 more move up actions than move down actions, although this is not accurate either
%% %%%% as move down actions that take chicken below the screen count as no-ops too ..
%% $\IW(1)$ does not have this limitation and is best in \Freeway.

@@

\noindent \textbf{Acknowledgments.} We thank the people involved in
the creation and maintenance of the Arcade Learning Environment that
have made this work possible. The research by Nir Lipovetzky is partially
funded by the Australian Research Council Linkage grant LP11010015.
